---
title: "Gene Set Libraries"
author: "Maintained by YuLab-SMU"
output:
  prettydoc::html_pretty:
    theme: cayman
    highlight: github
  pdf_document:
    toc: true
---



```{r echo=FALSE, results="hide", message=FALSE}
library(yulab.utils)
library(gson)
library(clusterProfiler)
library(kableExtra)
```

The `gson` file format was proposed in the CRAN R package `r CRANpkg("gson")`. 
This format is designed to store gene sets with related information (*e.g.*, gene set name, version, species, etc.). 

The `r CRANpkg("gson")` package provides a set of utilities to read, write and process `gson` files.

Here is a collection of gene sets stored in `gson` format:



```{r echo=FALSE}
# build gsonDf
files <- list.files(pattern = "\\.gson$")
fileDownload <- paste0("https://yulab-smu.top/gson-files/", files)
#icon <- paste0(icons::icon_style(icons::fontawesome("download"), fill='red'))
#icon <- paste0(icons::fontawesome("download", style = "solid"))
icon <- 'img/download-solid.svg'
#url <- glue::glue("[![]({icon})]({fileDownload})")
url <- glue::glue('[<img src="{icon}" width="30">]({fileDownload})')

library(gson)
gs <- lapply(files, read.gson)
geneSet <- vapply(gs, function(x) x@gsname, character(1))
species <- vapply(gs, function(x) x@species, character(1))
ver <- vapply(gs, function(x) {
    res <- x@version
    if (length(res) == 0) {
        return(" ")
    }
    return(res)
}, character(1))

ngs <- vapply(gs, function(x) length(unique(x@gsid2gene$gsid)), numeric(1))
ng <- vapply(gs, function(x) length(unique(x@gsid2gene$gene)), numeric(1))

d <- data.frame(`Gene Set` = geneSet,
                Terms = ngs,
                `Gene Coverage` = ng,
                Species = species,
                Version = ver,
                URL = url)

knitr::kable(d)

#library(kableExtra)
#d %>%
#  knitr::kable() #%>%
  #kable_styling()
  #kable_material(c("striped", "hover")) %>%
```





Users can download the file and use it as background annotation in `r Biocpkg("clusterProfiler")` package to perform enrichment analysis. 
 

```{r}
library(gson)
kegg <- read.gson("KEGG_human.gson")

library(clusterProfiler)
data(geneList, package="DOSE")
de = names(geneList)[abs(geneList) > 2] # select fold change > 2 as DEG
k1 <- enricher(de, gson=kegg)
head(k1, 2)

k2 <- GSEA(geneList, gson=kegg, verbose=FALSE)
head(k2, 2)
```


You are welcome to share your gson file with the community. Please create a [pull
request](https://github.com/YuLab-SMU/gson-files/pulls). 


